---
title: "Trade Statistics"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    social: [ "twitter", "facebook", "menu"]
    source_code: embed
    runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(knitr)
library(DT)
library(rpivotTable)
library(ggplot2)
library(plotly)
library(dplyr)
library(openintro)
library(highcharter)
library(ggvis)
library(gganimate)
library(shiny)
library(RMySQL)
```


```{r}
trades <- read.csv("c:/RTraining/Trade/data/secure/trades.csv")
topten <- read.csv("c:/RTraining/Trade/data/secure/tab.csv")


mydb <- dbConnect(MySQL(), user='root', password='', dbname='trade', host='localhost')

```

```{r}
mycolors <- c("blue", "#FFC125", "darkgreen", "darkorange")
```

Dashboard
=====================================

Row
-------------------------------------

### Goods Import/Export Analysis

```{r}
valueBox(paste("Current Month"),
         color = "warning")
```

### Total Current Month Import )

```{r}
rs4 = dbSendQuery(mydb, "SELECT sum(`Stat..Value`) as TotalImport_current from curtrade WHERE curtrade.CP4 = '4000' or curtrade.CP4 = '4071'  or curtrade.CP4 = '7100'")
    
    tab4 <- fetch(rs4, n=-1)
    import_cur <- ceiling(tab4$TotalImport_current/1000000)

valueBox(import_cur,
         icon = 'fa-arrow-left',
         color = "blue")

```

### Total Current Month Export

```{r}

rs3 = dbSendQuery(mydb, "SELECT sum(`Stat..Value`) as TotalExport_current from curtrade WHERE curtrade.CP4 = '1000' or curtrade.CP4 = '1021' or curtrade.CP4 = '1022' or curtrade.CP4 = '3071'")
    
    tab3 <- fetch(rs3, n=-1)
    
    export_cur <- ceiling(tab3$TotalExport_current/1000000)
    
    valueBox(export_cur,
         icon = 'fa-arrow-right',
         color = "green")


```

### Current Month Trade Balance

```{r}

tradebalance_cur <- export_cur - import_cur

valueBox(tradebalance_cur,
         icon = 'fa-balance-scale',
         color = "red")
```

Row
-------------------------------------

### Services Import/Export Analysis

```{r}
valueBox(paste("2019"),
         color = "warning")
```

### Total Services Export (2019 )

```{r}

service_export = dbSendQuery(mydb, "SELECT ServiceExport FROM tblservices WHERE Year ='2019'")
    
    servicesExport <- fetch(service_export, n=-1)
    serviceExport <- ceiling(servicesExport$ServiceExport)


valueBox(serviceExport,
         icon = 'fa-arrow-left',
         color = "blue")
```

### Total Services Import (2019)

```{r}
service_import = dbSendQuery(mydb, "SELECT ServiceImport FROM tblservices WHERE Year ='2019'")
    
    servicesImport <- fetch(service_import, n=-1)
    serviceImport <- ceiling(servicesImport$ServiceImport)

valueBox(serviceImport,
         icon = 'fa-arrow-right',
         color = "green")
```

### Services Trade Balance (2019)

```{r}

serviceBalance <- serviceExport - serviceImport

valueBox(serviceBalance,
         icon = 'fa-balance-scale',
         color = "red")
```


Row
-------------------------------

### Goods Export by Year

```{r}
p1 <- trades %>%
         group_by(Year) %>%
         summarise(Export = sum(Export_mil)) %>%
         plot_ly(x = ~Year,
                 y = ~Export,
                 color = "blue",
                 type = 'bar') %>%
layout(xaxis = list(title = "Year"),
yaxis = list(title = 'Export'))
p1
```

### Goods Import by Year

```{r}
p2 <- trades %>%
         group_by(Year) %>%
         summarise(Import = sum(Import_mil)) %>%
         plot_ly(x = ~Year,
                 y = ~Import,
                 color = "green",
                 type = 'bar') %>%
layout(xaxis = list(title = "Year"),
yaxis = list(title = 'Import'))
p2
```

### Goods Trade Balance by Year

```{r}
p3 <- trades %>%
         group_by(Year) %>%
         summarise(Balance = sum(TradeDef_mil)) %>%
         plot_ly(x = ~Year,
                 y = ~Balance,
                 color = "red",
                 type = 'bar') %>%
layout(xaxis = list(title = "Year"),
yaxis = list(title = 'Balance'))
p3
```

Row
------------------------------------

### Services Export by Year

```{r}
s2 <- trades %>%
         group_by(Year) %>%
         summarise(Export = sum(Service_exp)) %>%
         plot_ly(x = ~Year,
                 y = ~Export,
                 color = "green",
                 type = 'bar') %>%
layout(xaxis = list(title = "Year"),
yaxis = list(title = 'Export'))
s2
```

### Services Import by Year

```{r}
s1 <- trades %>%
         group_by(Year) %>%
         summarise(Import = sum(Service_imp)) %>%
         plot_ly(x = ~Year,
                 y = ~Import,
                 color = "blue",
                 type = 'bar') %>%
layout(xaxis = list(title = "Year"),
yaxis = list(title = 'Import'))
s1
```

### Services Trade Balance by Year

```{r}
s3 <- trades %>%
         group_by(Year) %>%
         summarise(Balance = sum(Service_bal)) %>%
         plot_ly(x = ~Year,
                 y = ~Balance,
                 color = "red",
                 type = 'bar') %>%
layout(xaxis = list(title = "Year"),
yaxis = list(title = 'Balance'))
s3
```

Row
---------------------------------
### Trade Trends in line graphs

```{r}
goodsExp <- dbSendQuery(mydb, "SELECT Year(Date) as Year, round(sum(Value)/1000000) as Totex FROM `historical_export_99_19` Where `Procedure`   = '1000' or `procedure` = '1021' or `Procedure` = '1022' or `Procedure` = '3071' Group By Year")
    goodsExport <- fetch(goodsExp, n=-1)

goodsImp <- dbSendQuery(mydb, "SELECT Year(Date) as Year, round(sum(Value)/1000000) as Totim FROM `historical_import_99_19` Where `Procedure`   = '4000' or `procedure` = '4021' or `Procedure` = '7100' Group By Year")
    goodsImport <- fetch(goodsImp, n=-1)
    
trade_combine_summary <- merge(goodsExport, goodsImport, by="Year" )
    trade_combine_summary$tbalance <- trade_combine_summary$Totex - trade_combine_summary$Totim  

```





Row
--------------------------------
### Trade Activities with 10 top countries


```{r}

imp <- read.csv("c:/RTraining/Trade/data/secure/country_year_import.csv")

imp %>% 
  mutate_if(is.factor, as.character)%>%
  mutate_if(is.integer, as.numeric) -> imp 

imp_formatted <- imp %>%
  group_by(Year) %>%
  mutate(rank = rank(-Value_mill),
         Value_rel = Value_mill/Value_mill[rank==1],
         Value_lbl = paste0(" ",round(Value_mill))) %>%
  group_by(CTY_Name) %>% 
  filter(rank <=10) %>%
  ungroup()

anim_imp <- ggplot(imp_formatted, aes(rank, group = CTY_Name, 
                                  fill = as.factor(CTY_Name), color = as.factor(CTY_Name))) +
  geom_tile(aes(y = Value_mill/2,
                height = Value_mill,
                width = 0.9), alpha = 0.8, color = NA) +
  geom_text(aes(y = 0, label = paste(CTY_Name, " ")), vjust = 0.2, hjust = 1) +
  geom_text(aes(y=Value_mill,label = Value_lbl, hjust=0)) +
  coord_flip(clip = "off", expand = FALSE) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_reverse() +
  guides(color = FALSE, fill = FALSE) +
  theme(axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position="none",
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x = element_line( size=.1, color="grey" ),
        panel.grid.minor.x = element_line( size=.1, color="grey" ),
        plot.title=element_text(size=25, hjust=0.5, face="bold", colour="grey", vjust=-1),
        plot.subtitle=element_text(size=18, hjust=0.5, face="italic", color="grey"),
        plot.caption =element_text(size=8, hjust=0.5, face="italic", color="grey"),
        plot.background=element_blank(),
        plot.margin = margin(2,2, 2, 4, "cm")) +
  transition_states(Year, transition_length = 4, state_length = 1, wrap = FALSE) +
  view_follow(fixed_x = TRUE)  +
  labs(title = 'Import per Year : {closest_state}',  
       subtitle  =  "Top 10 Countries Vanuatu has imported from",
       caption  = "Import Value in Billions of Vatu | Data Source: Asycuda")

animate(anim_imp, 200)
```

### Top 10 Export Countries

```{r}
exp <- read.csv("c:/RTraining/Trade/data/secure/country_year_export.csv")

exp %>% 
  mutate_if(is.factor, as.character)%>%
  mutate_if(is.integer, as.numeric) -> exp 

exp_formatted <- exp %>%
  group_by(Year) %>%
  mutate(rank = rank(-Value_mill),
         Value_rel = Value_mill/Value_mill[rank==1],
         Value_lbl = paste0(" ",round(Value_mill))) %>%
  group_by(CTY_Name) %>% 
  filter(rank <=10) %>%
  ungroup()


anim_exp <- ggplot(exp_formatted, aes(rank, group = CTY_Name, 
                                      fill = as.factor(CTY_Name), color = as.factor(CTY_Name))) +
  geom_tile(aes(y = Value_mill/2,
                height = Value_mill,
                width = 0.9), alpha = 0.8, color = NA) +
  geom_text(aes(y = 0, label = paste(CTY_Name, " ")), vjust = 0.2, hjust = 1) +
  geom_text(aes(y=Value_mill,label = Value_lbl, hjust=0)) +
  coord_flip(clip = "off", expand = FALSE) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_reverse() +
  guides(color = FALSE, fill = FALSE) +
  theme(axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position="none",
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x = element_line( size=.1, color="grey" ),
        panel.grid.minor.x = element_line( size=.1, color="grey" ),
        plot.title=element_text(size=25, hjust=0.5, face="bold", colour="grey", vjust=-1),
        plot.subtitle=element_text(size=18, hjust=0.5, face="italic", color="grey"),
        plot.caption =element_text(size=8, hjust=0.5, face="italic", color="grey"),
        plot.background=element_blank(),
        plot.margin = margin(2,2, 2, 4, "cm")) +
  transition_states(Year, transition_length = 4, state_length = 1, wrap = FALSE) +
  view_follow(fixed_x = TRUE)  +
  labs(title = 'Export per Year : {closest_state}',  
       subtitle  =  "Top 10 Countries Vanuatu has exported to",
       caption  = "Export Value in Billions of Vatu | Data Source: Asycuda") 

animate(anim_exp, 200)
```

Trade Trends by Countries
========================================


Row
----------------------------------------

Column {data-width=200 .sidebar}
----------------------------------------

```{r}
country = dbSendQuery(mydb, "SELECT CNTY_CODE, CNTY_DESCR FROM tblcountry")
cnty <- fetch(country, n=-1)

selectInput(inputId = "selectedCountry", label = "Select Country", choices = cnty$CNTY_DESCR)

```

Column
------------------------------------
## Selected Country Trend

```{r}


```


Export Trends by Commodities
========================================


Row
----------------------------------------

Column {data-width=200 .sidebar}
----------------------------------------

```{r}
comm = dbSendQuery(mydb, "SELECT Principle_Exports FROM tarif_r GROUP BY Principle_Exports")
pExport <- fetch(comm, n=-1)

selectInput(inputId = "selectedPrincipleExport", label = "Select Principle Export", choices = pExport$Principle_Exports)

```





Import Trends by Goods
========================================


Row
----------------------------------------

Column {data-width=200 .sidebar}
----------------------------------------


Download dataset 
========================================


Row
----------------------------------------

Column {data-width=200 .sidebar}
----------------------------------------




