---
title: "VNSO Trade Dashboard"
output: 
  flexdashboard::flex_dashboard:
    social: ["twitter", "facebook", "menu"]
    orientation: rows
    theme: cosmo
    vertical_layout: scroll
runtime: shiny
params:
  build_animations: TRUE
---

```{r setup, include=FALSE, warning=FALSE}
# Load libraries
library(flexdashboard)
library(tidyverse)
library(DT)
library(dplyr)
library(gganimate)
library(plotly)
library(knitr)
library(shiny)
library(formattable)
library(rgdal) # Reading and writing shape files
library(sp) # Working with spatial data
library(rmapshaper) # Simplifying polygons
library(leaflet) # Interactive mapping
library(countrycode) # Get country codes

# Note repository location
repository <- dirname(getwd()) # Note that to run code inline you'll need to set session working directory to source file location using Session in toolbar above
```

```{r Load files}
toptenexport <- read.csv(file.path(repository, "data", "secure", "tab.csv"))
countryYearExport <- read.csv(file.path(repository, "data", "secure", "country_export.csv"))
curtrade <- read.csv(file.path(repository, "data", "secure", "curtrade.csv"))
exp <- read.csv(file.path(repository, "data", "secure", "country_year_export.csv"))
imp <- read.csv(file.path(repository, "data", "secure", "country_year_import.csv"))
services <- read.csv(file.path(repository, "data", "secure", "tblservices.csv"))
trades <- read.csv(file.path(repository, "data", "secure", "trades.csv"))
tradenobalances <- read.csv(file.path(repository, "data", "secure", "trade.csv"))
#trade_ports <- read.csv(file.path(repository, "data", "secure", "trades_billions.csv"))



#************************** Create Summary Import historical *********************
#importHistory <- read.csv(file.path(repository, "data", "secure", "tradeStats_historic_IMPORTS_14-09-20.csv"))
#importHistorysummary <- importHistory %>%
#  group_by(Year, Month, Procedure, CTY_Origin, HS) %>%
#  summarise(Weight_Sum = sum(Weight),
#            Value_Sum = sum(Value), .groups="keep") %>%
#  ungroup()

#write.csv(importHistorysummary, "c:/temp/importsummaryhistory.csv")

importMajor <- read.csv(file.path(repository, "data", "open", "MAJOR_IMPORTS.csv"))
importHistorysummary <- read.csv(file.path(repository, "data", "secure", "imports_history_summary.csv"))



```

Dashboard
==================================================================

```{r Calculate current month import and export}
export <- curtrade %>%
  filter(CP4 %in% c(1000, 1021, 1022, 3071)) %>%
  summarise("Totex"=round(sum(Stat..Value, na.rm = TRUE)/1000000))

import <- curtrade %>%
  filter(CP4 %in% c(4000, 4071, 7100)) %>%
  summarise("Totim"=round(sum(Stat..Value, na.rm = TRUE)/1000000))

curtradebalance <- export - import

```

column {data-width=400}
------------------------------------------------------------------
### Total Current Month Export (Millions)

```{r Current trade moments}

valueBox(comma(export, digits = 0),
         icon = 'fa-balance-scale',
         color = "green")
```

### Total Current Month Import (Millions)
```{r Current trade imports}
valueBox(comma(import, digits = 0),
         icon = 'fa-balance-scale',
         color = "blue")

```

### Total Current Month Trade Balance (Millions)
```{r Current Month trade balance}
valueBox(comma(curtradebalance, digits = 0),
         icon = 'fa-balance-scale',
         color = "red")

```


```{r Calculate current year services import and export}
valueInterested <- services[services$year == 2019, ]

exportServices <- valueInterested$export
importServices <- valueInterested$import
serviceBalance <- exportServices - importServices

```


column {data-width=400}
------------------------------------------------------------------
### Services Export - 2019
```{r}
valueBox(comma(exportServices, digits = 0),
         icon = 'fa-balance-scale',
         color = "green")
```


### Services Import - 2019
```{r}

valueBox(comma(importServices, digits = 0),
         icon = 'fa-balance-scale',
         color = "blue")
```


### Services Trade Balance - 2019
```{r}

valueBox(comma(serviceBalance, digits = 0),
         icon = 'fa-balance-scale',
         color = "green")
```

Column {data-width=400}
-----------------------------------------------------------------------

### Trades in Goods and Services by Year (Billions)

```{r goods exports/imports/balance by year interactive plot}
trades %>%
  group_by(Year) %>%
  summarise(Export = sum(Export_Goods_RBV),
            Import = sum(Import_Goods_RBV),
            SExport = sum(Export_Services_RBV),
            SImport = sum(Import_Services_RBV), .groups="keep") %>%
  ungroup() %>%
  
  plot_ly(x = ~Year,
          y = ~Export,
          color = "blue",
          type = "scatter",
          mode = "lines+markers",
          name = "Export of Goods") %>%
  
  add_trace(x = ~Year,
            y = ~Import,
            color = "red",
            name = "Import of Goods") %>%
  
  add_trace(x = ~Year,
            y = ~SExport,
            color = "green",
            name = "Export of Services") %>%
  
  add_trace(x=~Year,
            y = ~SImport,
            color = "yellow",
            name = "Import of Services") %>%
  
  layout(xaxis = list(title = "Year"), yaxis = list(title = 'Value (Billion(Vatu))'))
```


### Trade Balances by Year (Billions)

```{r services exports/imports/balance by year interactive plot}
trades %>%
  group_by(Year) %>%
  summarise(GBalance = sum(TradeDef_RBV),
            SBalance = sum(Service_bal_RBV),
            TBalance = sum(Balance_RBV), .groups="keep") %>%
  ungroup() %>%
   
  plot_ly(x = ~Year,
          y = ~GBalance,
          color = "blue",
          type = "scatter", 
          mode = "lines+markers",
          name = "Goods Balance") %>%
  
  add_trace(x = ~Year,
            y = ~SBalance,
            color = "red",
            name = "Services Balance") %>%
  
  add_trace(x = ~Year,
            y = ~TBalance,
            color = "green",
            name = "Balance of External Goods & Services") %>%
  
  layout(xaxis = list(title = "Year"), yaxis = list(title = 'Value (Billion(Vatu))'))
```

Column
---------------------------------------------------------------------
### Trade in Goods through Port Vila Ports

```{r Trade in goods by Port}
trade_ports %>%
  group_by(Year) %>%
  summarise(GExport_Lug = sum(Lug_Export),
            GImport_Lug = sum(Lug_Import),
            GExport_Vil = sum(Vil_Export),
            GImport_Vil = sum(Vil_Import),
            POrts_Balance = sum(Balance), .groups="keep") %>%
  ungroup() %>%
   
  plot_ly(x = ~Year,
          y = ~GExport_Lug,
          color = "blue",
          type = "scatter", 
          mode = "lines+markers",
          name = "Luganville Goods Export") %>%
  
  add_trace(x = ~Year,
            y = ~GImport_Lug,
            color = "red",
            name = "Luganville Goods Import") %>%
  
  add_trace(x = ~Year,
            y = ~GExport_Vil,
            color = "green",
            type = "scatter",
            mode = "lines+markers",
            name = "Port Vila Goods Export") %>%
  
   add_trace(x = ~Year,
            y = ~GImport_Vil,
            color = "purple",
            type = "scatter",
            mode = "lines+markers",
            name = "Port Vila Goods Import") %>%
  
  layout(xaxis = list(title = "Year"), yaxis = list(title = 'Value (Billion (vatu)'))

```


### Trade Balance in Goods by Port

```{r Trade in goods Balances by Port}
trade_ports %>%
  group_by(Year) %>%
  summarise(GLug_Bal = sum(Lug_Bal),
            GVil_Bal = sum(Vil_Bal), .groups = "keep") %>%
  ungroup() %>%
   
  plot_ly(x = ~Year,
          y = ~GLug_Bal,
          color = "blue",
          type = "scatter", 
          mode = "lines+markers",
          name = "Luganville Trade Balance") %>%
  
  add_trace(x = ~Year,
            y = ~GVil_Bal,
            color = "red",
            name = "Port Vila Trade Balance") %>%
  
  
  layout(xaxis = list(title = "Year"), yaxis = list(title = 'Value (Billion (vatu)'))

```






Column {data-width=350}
-----------------------------------------------------------------------

### Trade Import Activities with 10 top countries overtime

```{r animated graphic for top 10 countries by year for imports, eval = params$build_animations}

imp %>% 
  mutate_if(is.factor, as.character) %>%
  mutate_if(is.integer, as.numeric) -> imp 

imp_formatted <- imp %>%
  group_by(Year) %>%
  mutate(rank = rank(-Value_mill),
         Value_rel = Value_mill/Value_mill[rank==1],
         Value_lbl = paste0(" ",round(Value_mill))) %>%
  group_by(CTY_Name) %>% 
  filter(rank <=10) %>%
  ungroup()

anim_imp <- ggplot(imp_formatted, 
                   aes(rank, group = CTY_Name, fill = as.factor(CTY_Name), color = as.factor(CTY_Name))) +
    
  geom_tile(aes(y = Value_mill/2, height = Value_mill, width = 0.9),
            alpha = 0.8, color = NA) +
  
  geom_text(aes(y = 0, label = paste(CTY_Name, " ")), vjust = 0.2, hjust = 1) +
  geom_text(aes(y=Value_mill,label = Value_lbl, hjust=0)) +
  
  coord_flip(clip = "off", expand = FALSE) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_reverse() +
  guides(color = FALSE, fill = FALSE) +
  
  theme(axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position="none",
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x = element_line( size=.1, color="grey" ),
        panel.grid.minor.x = element_line( size=.1, color="grey" ),
        plot.title=element_text(size=25, hjust=0.5, face="bold", colour="grey", vjust=-1),
        plot.subtitle=element_text(size=18, hjust=0.5, face="italic", color="grey"),
        plot.caption =element_text(size=8, hjust=0.5, face="italic", color="grey"),
        plot.background=element_blank(),
        plot.margin = margin(2,2, 2, 4, "cm")) +
  
  transition_states(Year, transition_length = 2, state_length = 1, wrap = FALSE) +
  
  view_follow(fixed_x = TRUE)  +
  
  labs(title = 'Import per Year : {closest_state}',  
       subtitle  =  "Top 10 Countries Vanuatu has imported from",
       caption  = "Import Value in Billions of Vatu | Data Source: Asycuda")

animate(anim_imp, nframes = 100, fps = 5)
```


### Trade Export Activities with 10 top countries overtime

```{r animated graphic for top 10 countries by year for exports, eval = params$build_animations}
exp %>%
  mutate_if(is.factor, as.character)%>%
  mutate_if(is.integer, as.numeric) -> exp 

exp_formatted <- exp %>%
  group_by(Year) %>%
  mutate(rank = rank(-Value_mill),
         Value_rel = Value_mill/Value_mill[rank==1],
         Value_lbl = paste0(" ",round(Value_mill))) %>%
  group_by(CTY_Name) %>% 
  filter(rank <=10) %>%
  ungroup()


anim_exp <- ggplot(exp_formatted, 
                   aes(rank, group = CTY_Name, fill = as.factor(CTY_Name), color = as.factor(CTY_Name))) +
  
  geom_tile(aes(y = Value_mill/2, height = Value_mill, width = 0.9),
            alpha = 0.8, color = NA) +
  
  geom_text(aes(y = 0, label = paste(CTY_Name, " ")), vjust = 0.2, hjust = 1) +
  geom_text(aes(y=Value_mill,label = Value_lbl, hjust=0)) +
  
  coord_flip(clip = "off", expand = FALSE) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_reverse() +
  guides(color = FALSE, fill = FALSE) +
  
  theme(axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position="none",
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x = element_line( size=.1, color="grey" ),
        panel.grid.minor.x = element_line( size=.1, color="grey" ),
        plot.title=element_text(size=25, hjust=0.5, face="bold", colour="grey", vjust=-1),
        plot.subtitle=element_text(size=18, hjust=0.5, face="italic", color="grey"),
        plot.caption =element_text(size=8, hjust=0.5, face="italic", color="grey"),
        plot.background=element_blank(),
        plot.margin = margin(2,2, 2, 4, "cm")) +
  
  transition_states(Year, transition_length = 4, state_length = 1, wrap = FALSE) +
  
  view_follow(fixed_x = TRUE)  +
  
  labs(title = 'Export per Year : {closest_state}',  
       subtitle  =  "Top 10 Countries Vanuatu has exported to",
       caption  = "Export Value in Billions of Vatu | Data Source: Asycuda") 

animate(anim_imp, nframes = 100, fps = 5)
```

Export by Commodity
=====================================

Column {.sidebar}
------------------------------------------------------------------
```{r Export by commodities, warning=FALSE}
startData <- toptenexport
GBChoices <- as.list(names(startData))
names(GBChoices) <- paste(names(startData),map(startData,~length(unique(.x))))

selectInput(inputId = "GB", label = "Group By", choices = GBChoices, selected = "Year 19")
selectInput(inputId = "Metric", label = "Metric", choices = names(select_if(startData,is.numeric)), selected = "Year")

updateData <- reactive(
  startData %>% 
    group_by_(input$GB) %>% 
    summarise_if(is.numeric, sum, na.rm=T)
  )
```

```{r Expot by countries}
countryExport <- countryYearExport
CEChoices <- as.list(names(countryExport))
names(CEChoices) <- paste(names(countryExport),map(countryExport,~length(unique(.x))))

updateCountryExport <- reactive(
  countryExport %>%
    group_by_(input$CE) %>% 
    summarise_if(is.numeric,sum,na.rm=T)
  )
```


Export by Commodity {data-navmenu="Trade in Export"}
=====================================

Column {.sidebar}
------------------------------------------------------------------
```{r}
selectInput(inputId = "GB", label = "Group By", choices = GBChoices)
selectInput(inputId = "Metric", label = "Metric", choices = names(select_if(startData,is.numeric)))
```

Column 
------------------------------------------------------------------

### Bar Plot of the selected commodity
```{r}
renderPlot({
  updateData() %>%
    ggplot(aes_string(x=input$GB, y=input$Metric, fill=input$GB)) +
      geom_col() +
      theme(panel.background = element_rect(fill = '#cccccc', colour = 'red'))
})
```

column
-------------------------------------------------
### Table didplay of the selected commodity 
```{r}
renderDT(
  updateData(), rownames = F, extensions = 'Buttons', filter="top", editable=T,
  options = list(
    dom = 'Blfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
    lengthMenu = list(c(10,50,100,-1),c(10,50,100,"All"))
  )
)
```

Export by Country {data-navmenu="Trade in Export"}
=====================================  

Column {.sidebar}
------------------------------------------------------------------
```{r}
countryExport <- countryYearExport
CEChoices <- as.list(names(countryExport))
names(CEChoices) <- paste(names(countryExport),map(countryExport,~length(unique(.x))))

selectInput(inputId = "CE",label = "Group By",choices = CEChoices, selected = "Year 19")
selectInput(inputId = "Metric_two",label = "Metric",choices = names(select_if(countryExport,is.numeric)), selected = "Year")

updateCountryExport <- reactive(
  countryExport %>%
    group_by_(input$CE) %>% 
    summarise_if(is.numeric,sum,na.rm=T)
  )
```

column
--------------------------------------
### Bar Plot of the export value by selected country
```{r}
renderPlot({
 updateCountryExport() %>% 
  ggplot(aes_string(x=input$CE,y=input$Metric_two,fill=input$CE)) +
  geom_col() +
  theme(panel.background = element_rect(fill = '#cccccc', colour = 'red'))
})
```

column
-----------------------------------------------------
### Table didplay of export value by selected country 
```{r}
renderDT(
  updateCountryExport(), rownames = F, extensions = 'Buttons', filter="top", editable=T,
  options = list(
    dom = 'Blfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
    lengthMenu = list(c(10,50,100,-1),c(10,50,100,"All"))
  )
)
```



Import by Goods {data-navmenu="Trade in Import"}
=====================================

```{r Import by goods of interest}
majorTen <- c("Rice"=10063000, "Wheat flour/ mesline flour"=11010000, "Cigarettes containing tobacco"=24022000, "Portland cement (excl. white)"=25232900, "Pharmaceutical products"=30049000, "Personal & Household effects"=98010000)

```

Column {.sidebar}
------------------------------------------------------------------
```{r}
selectInput(inputId = "MJTN",label = "Major 10 Goods",choices = majorTen)

```

```{r}

updateData_import_commodity <- reactive(
  importHistorysummary %>% 
    filter(HS == input$MJTN) %>%
    group_by(Year) %>%
    summarise(sum = Value_Sum, .groups = "drop")
  )

```

column
-------------------------------------------------
### Bar Plot of Import by Commodity

```{r}
renderPlot({
  updateData_import_commodity() %>% 
  ggplot(aes(x=Year, y=sum)) +
    geom_col() +
    theme(panel.background = element_rect(fill = '#cccccc', colour = 'red'))
})
```


column
-------------------------------------------------
### Table display of the selected commodity 
```{r}
renderDT(
  updateData_import_commodity(), rownames = F, extensions = 'Buttons', filter="top", editable=T,
  options = list(
    dom = 'Blfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
    lengthMenu = list(c(10,50,100,-1),c(10,50,100,"All"))
  )
)

```


Import by Country {data-navmenu="Trade in Import"}
===================================== 

Column {.sidebar}
------------------------------------------------------------------
```{r}
uniqueCountriesISO2 <- unique(importHistorysummary$CTY_Origin[is.na(importHistorysummary$CTY_Origin) == FALSE])
countries <- countrycode(uniqueCountriesISO2, origin = "iso2c", destination = "country.name")

countriesToSelectFrom <- uniqueCountriesISO2
names(countriesToSelectFrom) <- countries
countriesToSelectFrom <- countriesToSelectFrom[is.na(countries) == FALSE]

selectInput(inputId = "importCountry",label = "Metric", choices = countriesToSelectFrom)
```


```{r}

updateData_import_country <- reactive(
  importHistorysummary %>% 
    filter(CTY_Origin == input$importCountry) %>%
    group_by(Year) %>%
    summarise(sum = Value_Sum, .groups = "drop")
)

```


column
-------------------------------------------------
### Bar Plot of Import by Country

```{r}
renderPlot({
 updateData_import_country() %>%
  ggplot(aes(x=Year, y=sum)) +
  geom_col() +
  theme(panel.background = element_rect(fill = '#cccccc', colour = 'red'))
})
```

column
-------------------------------------------------
### Table display of the selected country 
```{r}
renderDT(
  updateData_import_country(), rownames = F, extensions = 'Buttons', filter="top", editable=T,
  options = list(
    dom = 'Blfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
    lengthMenu = list(c(10,50,100,-1),c(10,50,100,"All"))
  )
)

```


Trade through Maps
===============================================

```{r load world map, include = FALSE}
# Download world country boundaries from: https://www.naturalearthdata.com/downloads/50m-cultural-vectors/50m-admin-0-countries-2/ (https://www.naturalearthdata.com/http//www.naturalearthdata.com/download/50m/cultural/ne_50m_admin_0_countries.zip)
# Get a map of the world as as a spatial polygons object
shapeFile <- file.path(repository, "data", "secure", "ne_50m_admin_0_countries", "ne_50m_admin_0_countries.shp")
world <- readOGR(shapeFile, verbose = FALSE)

# Simplify the polygons
worldSimple <- ms_simplify(world, keep = 0.1)

# Keep necessary columns for joining
worldSimple@data <- worldSimple@data[, c("NAME", "CONTINENT", "ISO_A2", "ISO_A3")]
```

```{r join world with trades data}
# Join in the imports data
worldImports <- merge(worldSimple, imp[imp$Year == 2018, c("CTY_ID", "Value_mill")], by.x = "ISO_A2", by.y = "CTY_ID")

# Join in the exports
worldImportsAndExports <- merge(worldImports, exp[exp$Year == 2018, c("CNTY_CODE", "Value_mill")],
                                by.x = "ISO_A2", by.y = "CNTY_CODE",
                                suffixes = c("_Imports", "_Exports"))
```

```{r build interactive map}
# Create a colour palette
valueRange <- range(c(worldImportsAndExports$Value_mill_Imports, 
                      worldImportsAndExports$Value_mill_Exports), na.rm = TRUE)
colours <- c("blue", "white", "red")
colourPalette <- colorNumeric(colours, valueRange)
colourPaletteWithoutNA <- colorNumeric(colours, valueRange, na.color=rgb(0,0,0,0))

# Create labels that will appear on hover
labels <- paste0("<strong>", worldImportsAndExports$NAME, "</strong>",
                 "<br>Import value: ", round(worldImportsAndExports$Value_mill_Imports, digits=2),
                    ifelse(is.na(worldImportsAndExports$Value_mill_Imports), "", " million"),
                 "<br>Export value: ", round(worldImportsAndExports$Value_mill_Exports, digits=2),
                 ifelse(is.na(worldImportsAndExports$Value_mill_Exports), "", " million")) %>% lapply(htmltools::HTML)

# Create an interactive map
leaflet(worldImportsAndExports, options = leafletOptions(crs = leafletCRS(crsClass = "L.CRS.EPSG4326"))) %>%

  setView(lng = 0, lat = 0, zoom = 1) %>%
  
  addPolygons(weight = 2, color = "white", dashArray = "3", fillOpacity = 0.5,
              fillColor = ~colourPalette(Value_mill_Imports),
              highlight = highlightOptions(weight = 5, color = "#666", dashArray = "",
                                           fillOpacity = 1, bringToFront = TRUE),
              label = labels,
              labelOptions = labelOptions(style = list("font-weight" = "normal", padding = "3px 8px"),
                                          textsize = "15px", direction = "auto"),
              group = "Imports") %>%

  addPolygons(weight = 2, color = "white", dashArray = "3", fillOpacity = 0.5,
              fillColor = ~colourPalette(Value_mill_Exports),
              highlight = highlightOptions(weight = 5, color = "#666", dashArray = "",
                                           fillOpacity = 1, bringToFront = TRUE),
              label = labels,
              labelOptions = labelOptions(style = list("font-weight" = "normal", padding = "3px 8px"),
                                          textsize = "15px", direction = "auto"),
              group = "Exports") %>%
  
  addLegend(values = valueRange, bins = seq(from = 0, to = signif(valueRange[2], 2), length.out = 5),
            pal = colourPaletteWithoutNA, title = "Value (millions)",
            position = "bottomleft", na.label = "") %>%
  
  addLayersControl(baseGroups = c("Imports", "Exports"), 
                   options = layersControlOptions(collapsed=FALSE)) %>%
  hideGroup(c("Exports"))
```
