---
title: "Exploring historic data"
author: "Vanuatu National Statistics Office"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: '2'
    toc_float: yes
params: 
  password: "ThisIsNotMyPassword"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r preparation, include=FALSE}

# Load the required libraries
library(RMySQL) # Interactiing with MySQL from R
library(knitr) # Nice tables
library(kableExtra) # Extra nice tables
library(basicPlotteR) # Progress bar
library(plotly) # Interactive graphs

# Load the general R functions
source("functions.R")
```

# Introduction

Here we document the exploration of the historic trade statistics data for Vanuatu. Our aim is to understand how the value of different imported and exported commodities has changed through time. Due to the large nature of the historic data, these data are stored on a [MySQL](https://www.mysql.com/) server. We will use the [RMySQL](https://cran.r-project.org/web/packages/RMySQL/RMySQL.pdf) R package to interact with the local MySQL server from R.

To set up the local MySQL server, a dump of the VNSO server was shared and imported via the [MySQL Workbench](https://www.mysql.com/products/workbench/) on my computer. I had previously install MySQL by following these instructions:
- Installing on linux systems ([link](https://dev.mysql.com/doc/refman/8.0/en/linux-installation.html)), with extra help [here](https://itsfoss.com/install-mysql-ubuntu/).
- Installing on mac: ([link](https://dev.mysql.com/doc/mysql-osx-excerpt/5.7/en/osx-installation.html)), also, I think it comes ready installed! See [here](https://www.thoughtco.com/installing-mysql-on-mac-2693866))
- Installing on Windows ([link](https://dev.mysql.com/downloads/installer/)), with extra help [here](https://www.wikihow.com/Install-the-MySQL-Database-Server-on-Your-Windows-PC)

# Connecting to the database

The MySQL server is running locally and we can connect to it from R with the following code:

```{r connection, eval=FALSE}
# Open a connection to a MySQL database
connection <- dbConnect(MySQL(), 
                            user='root', 
                            password=readline(prompt="Enter password: "), # Doing this as password never stored in easily accessible format now
                            dbname='vnso',
                            host='localhost')
```


```{r echo=FALSE}
# Open a connection to a MySQL database
connection <- dbConnect(MySQL(), 
                        user='JosephCrispell', 
                        password=params$password, # Doing this as password never stored in easily accessible format now
                        dbname='vnso',
                        host='localhost')
```

```{r store local csv, eval=FALSE, echo=FALSE}
# Get the historic data
historic <- dbGetQuery(conn=connection, statement="SELECT * FROM historical_export_99_19")

# Save the data as a standard csv
write.table(historic, file=file.path("..", "data", "secure", "tradeStats_historic_14-09-20.csv"),
            sep=",", row.names=FALSE, quote=TRUE)
```
 
On our local MySQL database there is a single table `historical_export_99_19`, which looks like this:

```{r quick look at historic data, echo=FALSE}

# Extract a small subset of historic data
historic <- dbGetQuery(conn=connection, statement="SELECT * FROM historical_export_99_19 LIMIT 25")

# View the table
prettyTable(historic)
```

# Summarise the `Value` distribution for each HS code

Each commodity in the historic data is identified by an 8 digit **HS** (Harmonized System) code. We are aiming to define the expected `Value` distribution for each unique **HS** code through time. We can start with some simple summary statistics:

```{r summarise value column}
# Summarise the value column for each commodity for each year
commoditySummary <- dbGetQuery(conn=connection, statement="SELECT HS, YEAR, AVG(Value) AS average, MAX(Value) AS maximum, MIN(Value) AS minimum, COUNT(Value) AS count, SUM(Value) AS sum, STDDEV_SAMP(Value) AS sd, VAR_SAMP(Value) as variance FROM historical_export_99_19 GROUP BY HS, YEAR")

# Remove NA values
commoditySummary <- commoditySummary[is.na(commoditySummary$HS) == FALSE, ]

# Sort the commodity by HS code and year
commoditySummary <- commoditySummary[order(commoditySummary$HS, commoditySummary$YEAR), ]

# View the table
prettyTable(commoditySummary[1:25, ])
```

<br>
With the above data we can see that there are `R length(unique(commoditySummary$HS))` *HS* codes. There's one code that dates back to 1900!
<br>

```{r 1900, echo=FALSE}
# Get the information for the HS code with information dating back to 1900
old <- commoditySummary[commoditySummary$HS == "84715000", ]

# View the data
old[order(old$YEAR), ]
```

# Closing a connection to the `MySQL` server

To wrap up, we can close our connection to the `MySQL` server using the following code:
```{r warning=FALSE}
dbDisconnect(conn=connection)
```

